Contextuality and the Born rule in fuzzy quantum theories

### 2017-05-10 prop_letter.tex
From: Hanson, Andrew J. Sent: Wednesday, May 10, 2017 11:27 To: Tai, Yu-Tsung <yuttai@indiana.edu> Cc: Sabry, Amr A. <sabry@indiana.edu>; Ortiz, Gerardo <ortizg@indiana.edu> Subject: Here is the first draft intro text
Here is our starting draft of the intro page text:
   -- the last paragraphs in particular depend a lot on opinions of what
we should emphasize, and so are heavily open to discussion, revision,
and improvement.
    -Andy and Yu-Tsung
----------------------------------------------------

In this letter, we extend our previous work on the foundations of
computability in quantum mec

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Textbook quantum mechanics asserts that the probability of a measurement
can be computed from wave function of the system and the measuring
observable \emph{exactly} while real measurements must be finite precision.
Whether finite precision measurement in reality affect the idealized
quantum concept, e.g., contextuality~\cite{PhysRevLett.83.3751,Mermin1999,BarrettKent2004},
and the very foundation of quantum mechanics is still a problem. In
this letter, we model the finite precision quantum measurement by
quantum interval-valued probability measures (IVPMs) which provides
not only a uniform framework to understand Gleason's theorem~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}
and the Kochen-Specker theorem~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA,Griffiths2003}
but also some examples which cannot be explained by the Born rule~\cite{Born1983,peres1995quantum,544199,Jaeger2007}.

The tension between finite precision measurement and the foundation
of quantum mechanics sharpened, when the original definition of contextuality
based on infinitely precise numbers is proved not experimental verifiable~\cite{PhysRevLett.83.3751,Kent1999}.
After a long debate~\cite{Mermin1999,Peres2003,BarrettKent2004},
people are still trying to update the definition of contextuality
when they want to perform further experiments~\cite{Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}.

We found that finite precision measurement affects not only contextuality
but also the fundamental rule computing probability from a state.
In the early formulation of quantum mechanics, von Neumann essentially
formulated a quantum state as a density matrix while Mackey formulated
a quantum state as a quantum probability measure which are proved
by Gleason to be the essentially same as a density matrix~\cite{Varadarajan2008}.
Combining quantum probability measures and the classical IVPM~\cite{JamisonLodwick2004}
which could be used to formulate classical finite precision measurement,
we can define quantum IVPMs. Surprisingly, there is a quantum IVPM
which cannot be compatible with any density matrix according to the
Born rule. Similar to revising the definition of contextuality because
of finite precision measurement, the very basic definition of a quantum
state or how to compute the probability from a state might need to
be refined when we consider the finite precision measurement. This
can also provide a new quantum phenomena different from the classical
world, because IVPMs become classical so that the non-Born IVPMs disappears
when we restricting the domain of quantum IVPMs on a commuting sub-event
space. \newpage{}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this letter, we extend our previous work on the foundations of computability in quantum mechanics by exploring the application of interval-valued probability measures (IVPM) to achieving a coherent formulation of finite-resource quantum mechanics.

There are long-standing debates in the foundations of quantum mechanics regarding the tension between finite precision measurement and contextuality~\cite{Peres2003,BarrettKent2004}.  For example, Meyer~\cite{PhysRevLett.83.3751} and Mermin~\cite{Mermin1999} disagree on whether finite-precision measurement invalidate the Kochen-Specker theorem~\cite{kochenspecker1967,peres1995quantum}, and discussion continues on whether contextuality is experimentally verifiable~\cite{Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}. 
The
fundamental question is how to make mathematical theories of quantum mechanics that are intrinsically, rather than only implicitly, consistent with the resources available for the realistic accuracy of an actual measurement.

Our first observation is that essential objectives of our program to achieve a computable theory of quantum mechanics and finite-resource measurement, attempted previously with computable number systems~\cite{usat,geometry2013,DQT2014}, may be achievable by extending classical interval-valued probability measures~\cite{JamisonLodwick2004} to the quantum domain.  We begin by axiomatizing a quantum interval-valued probability measure (QIVPM) framework exploiting the fact that a density matrix approach can be considered as equivalent to a description of quantum states using probability measures~\cite{10.2307/2308516,Varadarajan2008}. %Mackey=10.2307 One result is that the expectation value of an observable with respect to a QIVPM can be defined to be consistent with the constraints on the expectation value of the same observable expected from continuous quantum mechanics.  We then examine Gleason's theorem~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum} and the Kochen-Specker theorem in the context of QIVPM's, and conclude that they correspond to the limit of an infinite number of precise intervals, and the set of crypto-deterministic intervals (the set $\{\imposs,\necess\}$), respectively.

We note, however, that in this approach there exist QIVPM's that do not correspond to a density matrix computable by the Born rule, so that Gleason's theorem cannot hold in general.  It thus remains possible that a refinement of the basic definition of a quantum state or the computation of probability from a state may be required when we rigorously consider the implications of finite precision measurement.
